---
title: "COVID-19 impacts on domestic and international students"
author: "Group 161 in Canvas; 490402742, 500459007, 500235825, 500625831"
subtitle: "Project 2"
date: "University of Sydney | DATA1001 | September 2020"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

<br>

# Executive Summary

This study examines the effect of COVID-19 on domestic and international students. Data reflects **high unemployment rates** and dependence on external financial support for both groups and reveals a **weak correlation** with hours studied. COVID-19 has negatively affected the student education but has not significantly influenced the consideration to defer.  

<br>

# Full Report

## Initial Data Analysis (IDA)

```{r, echo=F}
# Read in your survey data
survey = read.csv("realsurvey.csv")
```

### Source

Data was collected from a survey made by data science students from the University of Sydney using Google Forms. 

The survey was distributed to students across multiple universities via an EdStem forum and social media.


### Structure

Survey consists of 12 questions, recording 35 responses.


### Limitations

- Data was collected over a *short* period of time (1 week).
  - This may not be reflective of overall, year-length student experiences
  
<br>

- Sample size of 34 students may not be a realistic representations of thousands of university students

- Potential bias in survey results

- Responses from domestic students dominate over international students.
  -  Less international students being represented
  -  Possibly due to survey being inaccessible
  - Google is banned in mainland China, which is where most international students reside
  
<br>

- Questions based on assumptions and general averages (hours of study) is dependent on other factors and variables (such as amount of homework)
  


<br>

## Research Questions 

### 
How many hours do students work?
```{r, echo=F}
domw = read.csv("domw.csv")
test  = table(domw)
barplot(test, main = "Working hours for domestic students", xlab = "Number of students", ylab = "Hours worked", horiz = T, col = "purple")

intw = read.csv("intw.csv")
test  = table(intw)
barplot(test, main = "Working hours for international students", xlab = "Number of students", ylab = "Hours worked", horiz = T, col = "purple")
```

From the sample data, most domestic students are **currently unemployed**. The next highest being 5 hours a week and 21 hours a week. This trend continues with *international students* as most are unemployed as well.

<br>

###
Has student working hours increased or decreased due to COVID-19?
```{r, echo=F}
domr = read.csv("domr.csv")
test = table(domr)
library(RColorBrewer)
coul <- brewer.pal(5, "Set2")
barplot(test, main = "Changes in work for domestic students", col=coul)
```

```{r, echo=F}
intr = read.csv("intr.csv")
test = table(intr)
library(RColorBrewer)
coul <- brewer.pal(5, "Set2")
barplot(test, main = "Changes in work for international students", col=coul)
```

A higher proportion of international students have reported **decreased work hours** in comparison to domestic students which can suggest a *higher financial burden* experienced by international students. 

<br>

###
Does the financial support received due to COVID-19 differ between domestic and international students? 
```{r, echo=F}
domf = read.csv("domfin.csv")
tabel = table(domf)
library(RColorBrewer)
coul <- brewer.pal(5, "Set3")
barplot(tabel, main = "Domestic students financial support", names.arg=c("Govt help","Family/friends","Jobkeeper", "Uni support"), col=coul)

intf = read.csv("intfin.csv")
hell = table(intf)
library(RColorBrewer)
coul <- brewer.pal(5, "Set3")
barplot(hell, main = "International students financial support", names.arg=c("Family/friends","Jobkeeper", "Uni support"), col=coul)
```

For both domestic and international students, **financial support from family and friends** is the main form of aid received. 

COVID-19 supplements is also a widely received support option with domestic students. However, it is not present with international students as other countries/governments *do not* provide such aid.

<br>

###
Is there a relationship between the hours worked and the number of hours studied during COVID-19?

```{r, echo=F}
survey = read.csv("realsurvey.csv")

library(ggplot2)
ggplot(survey, aes(x = How.many.hours.of.work.do.you.work.a.week.during.COVID.19., y = How.many.hours.of.study.did.you.do.during.COVID.19.)) + geom_point() + geom_smooth(method = 'lm', se = F) + labs(x='Work (hrs)', y = 'Study (hrs)')

work = survey$How.many.hours.of.work.do.you.work.a.week.during.COVID.19.
study = survey$How.many.hours.of.study.did.you.do.during.COVID.19.


M = lm(study~work)
summary(M)
cor(work,study)

plot(work, M$residuals, abline(h=0))
```

This graph has a *very weak positive trendline* with a correlation coefficient of 0.25, suggesting a weak relationship between the number of hours worked with the hours studied. The linear equation is represented by study = 9.09 + 0.159 (work).

The data appears to be heteroscedastic hence why a linear model is not suitable but this might also be due to the lack of participants. Surprisingly, despite the weak relationship it suggests that the **more hours worked correlated with higher hours of study** which also highlights that time may not be the only major contributing factor to student studies. 

Other factors like work ethic, motivation and lifestyles and one's studyload may influence the hrs of study undertaken by students. 

<br>

###
Have the challenges of COVID-19 towards the quality of education received contributed to domestic and international student’s consideration of deferring their studies?
```{r, echo=F}
ddom = read.csv("ddom.csv")
domd = table(ddom)
library(RColorBrewer)
coul <- brewer.pal(5, "Set2")
barplot(domd, main = "COVID-19 positively/negatively impacted studies of domestic students", xlab = "Domestic students response", ylab = "Number of students", col=coul)


dint = read.csv("dint.csv")
intd = table(dint)
library(RColorBrewer)
coul <- brewer.pal(3, "Set2")
barplot(intd, main = "COVID-19 positively/negatively impacted studies of int. students", xlab = "International students response", ylab = "Number of students", col=coul)


sdom = read.csv("sdom.csv")
tabl = table(sdom)
barplot(tabl, main = "Effectiveness of online learning for domestic students", xlab = "Domestic students response", ylab = "Number of students", col = "orange")


sint = read.csv("sint.csv")
tabl = table(sint)
barplot(tabl, main = "Effectiveness of online learning for int. students", xlab = "International students response", ylab = "Number of students", col = "orange")
```

It is evident that the majority of both domestic and international students agree that COVID-19 has **negatively affected their education**. However, the data *differs* when considering the effectiveness of online learning during COVID-19. With domestic students, majority *agree* on a 6/10 rating, whilst for international students the rating *varies*, with 4/10, 6/10 and 8/10 being popular options. 

This could indicate that **COVID-19 has not necessarily contributed directly** to students deferring studies as there are some who do prefer online learning or feel a positive impact on their studies with their current circumstances.

<br>

###
Do international students consider deferring studies more than domestic students? If so, how come?
```{r, echo=F}
thoughtd = read.csv("thoughtd.csv")
domt = table(thoughtd)
library(RColorBrewer)
coul <- brewer.pal(5, "Set1")
barplot(domt, main = "Domestic students decision to defer studies", xlab = "Student response", ylab = "Number of students", horiz = T, col=coul)
```

```{r, echo=F}
thoughti = read.csv("thoughti.csv")
intt = table(thoughti)
library(RColorBrewer)
coul <- brewer.pal(5, "Set1")
barplot(intt, main = "International students decision to defer studies", xlab = "Student response", ylab = "Number of students", horiz = T, col=coul)
```

From the data collected, it would seem that **similar proportions** of both domestic and international students have considered deferring studies this year, although a higher portion of domestic studnets considered deferring in comparison to domestic students.

<br>


## Related Articles
International students have been reported to experience financial hardships during COVID-19 with many experiencing job cuts especially with being excluded from government help programs (Purtill, 2020). This added financial stress with a predominantly online learning environment has been shown to negatively impact student studies (Morris et al., 2020). 


<br>

## References

- Morris, A., Hastings, C., Wilson, S., Mitchell, E., Ramia, G., & Overgaard, C. (2020). The Experience of International Students Before and During COVID-19. Retrieved from Analysis & Policy Observatory: https://bit.ly/3kljJrO

- Purtill, J. (2020). ‘It has really f*****d me over’: The impact of COVID on young hospitality. Retrieved October 21, 2020, from  https://www.abc.net.au/triplej/programs/hack/covid-19-coronavirus-pandemic-hit-young-hospitality-workers-hard/12799238 


<br>

## Acknowledgments
Team consists of Ella, Keane, Leo and Yunhan. Group met on the 28th of September, around 1:30pm (GMT+11) via Facebook.

Ella contributed in the making of the survey through Google Forms and analysing results.

Keane coded the report in RStudio and assisted in making the survey.

Both Leo and Yunhan contributed by suggesting possible questions for the survey and distributing it to peers for data collection.








