---
title: "Newtown 2017 House Prices Report"
author: "**SID:** 500459007"
subtitle: "Project 1"
date: "University of Sydney  | DATA1001  |  September 2020"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

<br>

# Executive Summary

From the data analysis and report, it is evident that the price of properties is affected by different variables such as number of bedrooms, carspots etc.

Properties in Newtown are rather expensive (average price: $1.4 mil), with the cheapest being $370k and the most expensive being $3.15 mil.


<br>

# Full Report

## Initial Data Analysis (IDA)

The Newtown property prices dataset comes from data scraped by [domain](www.domain.com.au) in June 2017. 

<br>

### Source:

The data is sourced from [Domain](www.domain.com.au), a reputable and legitimate Australian digital property portal. 

In terms of the data collected, there is a possibility of:

- Human error during entering data in the system
- Miscalculations or misinterpretation of data 
- False or invalid data entry

### Structure

The data has 56 objects with 8 variables

```{r}
property = read.csv("http://www.maths.usyd.edu.au/u/UG/JM/DATA1001/r/current/data/NewtownJune2017.csv")
dim(property)
```

```{r, echo=F}
str(property)
```

```{r, echo=F}
count = table(property$Bedrooms)
barplot(count, main = "Bedrooms per property", xlab = "Bedrooms", ylab = "Properties")
```

```{r, echo=F}
bath = table(property$Bathrooms)
barplot(bath,main = "Bathrooms per property", xlab = "Bathrooms", ylab = "Properties")
```

```{r, echo=F}
spots = table(property$Carspots)
barplot(spots, main = "Carspots per property", xlab = "Carspots", ylab = "Properties")
```

```{r, echo=F}
sells = table(property$Sold)
plot(sells, main = "Property selling price", xlab = "Price ($)", ylab = "Properties sold")
```

<br>

### Limitations
```{r, echo=F}
dates = table(property$Date)
plot(dates, main = "Dates of property sold", xlab = "Date", ylab = "Properties sold")
```

- All the properties sold were **within 2 months** and **only within 1 suburb (Newtown)**. 
- **Small sample size** and the **timing of data collection** may not be an appropriate reflection of house prices *throughout* Sydney.
- **Confounding variables** (such as the property location) may also affect prices.


<br>


## Research Questions (given)

### What was the average price of a property?

Property average price: 
```{r, echo = F}
mean(property$Sold)
```

Property median price:
```{r, echo=F}
median(property$Sold)
```

Standard deviation of prices:
```{r,echo=F}
sd(property$Sold)
```

IQR of prices:
```{r, echo = F}
IQR(property$Sold)
```

<br>

### Does the amount of carspots affect the price?
```{r, echo = F}
library(ggplot2)
p = ggplot(property, aes(Sold, Carspots))
p + geom_point()
```

Mean of properties with 1 carspot:

```{r, echo=F}
mean(property$Sold[property$Carspots=="1"])
```

Mean of properties with 2 carspots:
```{r, echo=F}
mean(property$Sold[property$Carspots=="2"])
```

Mean of properties with no carspots:
```{r, echo=F}
mean(property$Sold[property$Carspots=="0"])
```


The amount of carspots **does seem to affect** the price of properties as the mean for properties with 2 carspots is *much higher* than for properties with 1 or no carspots. 

However, it is evident that most properties either have 1 or no carspots. This sample **may not be an accurate representation** of the number of carspots affecting the price.

```{r,echo=F}
spots = table(property$Carspots)
barplot(spots, main = "Number of carspots per property", xlab = "Carspots", ylab = "Properties")
```

<br>

###  What were the cheapest properties?

Properties sorted by price:
```{r, echo=F}
sold = property$Sold
sort(sold)
```

Cheapest properties
```{r, echo=F}
attach(property)
head(property[order(Sold), ][c(1, 6, 7)], 10)
```

Top 3 cheapest properties were:

- 514/88 King St, $370k
- 14/163 Australia St, $625k
- 1/60 Wilson St, $645k

<br>

### Assuming commission fees are around 2%, which agency earnt the most over that period? 

Agent sales:
```{r, message = FALSE, echo=F}
library(tidyverse)
property_grouped_by_agent <- group_by(property, Agent)
summarise(property_grouped_by_agent, earnings = sum(Sold)*0.02)
```
*RayWhite* is the agency with most earnings (347.62)

<br>


## Research Question (own)

### Does the number of bedrooms affect the price of properties?  

Average price with 1 bedroom:
```{r, echo=F}
mean(property$Sold[property$Bedrooms=="1"])
```

Average price with 2 bedrooms:
```{r, echo=F}
mean(property$Sold[property$Bedrooms=="2"])
```

Average price with 3 bedrooms:
```{r, echo=F}
mean(property$Sold[property$Bedrooms=="3"])
```

Average price with 4 bedrooms:
```{r, echo=F}
mean(property$Sold[property$Bedrooms=="4"])
```

Average price with 5 bedrooms:
```{r, echo=F}
mean(property$Sold[property$Bedrooms=="5"])
```  

<br>

Number of bedrooms vs Price
```{r, echo = F}
library(ggplot2)
p = ggplot(property, aes(Sold, Bedrooms))
p + geom_point()
```

Number of bedrooms in each property **affects the price**.

As  **bedrooms increase**,  average price **increases**. 

Average for 5 bedrooms is *lower* than average for 4 bedrooms as **only 1 property** has 5 bedrooms.

1 property has **no bedrooms**, possibly due to data misinterpretation or it is not for living (a garage).





